### 2015-06-03 doc/paper.tex
Fix some latex glossary instructions

% Dedicated to _why the lucky stiff, wherever on the net you may be
\documentclass[12pt]{article}
\usepackage{color}
\usepackage{hyperref}
\usepackage{mdwlist}
\usepackage{longtable}
\usepackage{ulem}
\usepackage{listings}
\usepackage[xindy]{glossaries}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphics}

@@

\normalem
\hypersetup{colorlinks=true, linkcolor=blue, filecolor=blue, pagecolor=blue, urlcolor=blue}
\makeglossary

@@

Creating an open implementation of a modern decompression algorithm could make diving much safer, and
provide divers with software that could be shown to be safe. This report outlines the design and implementation
of a proof of concept implementation of the VPM-B \newglossaryentry{VPM-B}{ description={Varying Permeability Model}}
algorithm, in the hopes that it can provide a base on 
which to build production quality software and hardware for an open dive computer.

@@

\subsubsection{VPM}
 The VPM algorithm starts by adopting a model of body tissues developed by Albert A. B\"uhlmann. This model
uses 16 tissue compartments to model the effect the dissolved gasses have on human tissues. During the
normal part of the dive, gas is added to the compartments\cite{Buhlmann84} (on-gassing\newglossaryentry{On-Gassing} {description={The process of adding inert gasses to body tissues}}
), and during the final ascent,
they're released from the tissue (off-gassing\newglossaryentry{Off-Gassing} { description={The process of releasing inert gasses from body tissues}}
). Many different compartments are needed because different
tissues absorb and release gasses at different rates (as evidenced by the Helium\_Half\_Time and
Nitrogen\_Half\_Time arrays in the program).

@@

To reduce the time decompressing without increasing the danger, the Critical Volume Algorithm is applied
during the ascent. The CVA\newglossaryentry{CVA} { description={Critical Volume Algorithm, the algorithm that allows the VPM algorithm to give decent decompression times.}}
can't reduce the time of a ``no bubble growth'' ascent, but it can limit
the growth of the bubbles to a safe amount. This is done by calculating a ``no bubble growth'' ascent,
and then applying the CVA iteratively until the ascent is judged safe.

@@

For this honours project, two open source implementations of the algorithm VPM-B algorithm were written. A version designed to be
usable in a desktop setting to serve as a reference implementation, and allow the user to generate decompression
tables, and a version able to work on an embedded computer, and do real time measurements of depth and time, to
build accurate models of the gasses in a divers system. Furthermore, these implementations are released under the
``Simplified BSD License'' \newglossaryentry{BSD License} { description={A family of software licenses that allow the use of the code in open source or proprietary software}}
, which allows them to be used in a proprietary or open source dive computer. While
this does allow a company to simply take the code, modify it, and use it in their product without
giving back the changes, it is an acceptable compromise to it not being used at all, because it is under
a more restrictive license.

@@

 Once the Python version was up and running, it was time to make many changes to the code to make it more
in line with the programming practices taught at Carleton University. The original program was written for FORTRAN 77,
and followed many of the practices common to FORTRAN\newglossaryentry{FORTRAN} { description={A programming language, popular with scientific researchers, but
not often taught to programmers these days}}
programs (GOTOs, global state used for all long term storage,
custom file formats for input data), which would be very hard for modern day programmers to follow. The program
was made object oriented\footnote{There are definite instances of the ``God Object'' anti-pattern, but at least the state is encapsulated, and in a position to be refactored}, by ensuring that all the global state was encapsulated in a DiveState object, and all functions that would modify this state belonged to the object.
The GOTOs were turned into normal loops with break statements, extremely long functions were broken into smaller
pieces, and repeating code was extracted to stand alone methods. Finally,  the input and output file formats were changed
to use JSON,\newglossaryentry{JSON} { description={JavaScript Object Notation, a lightweight data representation language
designed to be easier to write than XML}}
 as it is a standard format like XML\newglossaryentry{XML} { description={Extensible Markup Language, a language used
to structure data for easy computer reading and writing}}
, that is supported by just about every programming language \footnote{Though not FORTRAN, so a translator was written to let the original program load the new input files}.

 Now that the Python program was finished, the port to C\newglossaryentry{C} { description={A programming language, designed for small size, and fast execution, which is 
well suited for embedded systems}}
 was started. This was fairly straightforward, but it did involve
much more work to load JSON files than the Python version\footnote{The Python parser can turn it 
directly into hash-tables and arrays, while the C version
needs load values into the structures one at a time}. The port also showed a few bugs in the rarely run parts
of the Python program, that ended up being caught by the static type checks.

>>>>

% Dedicated to _why the lucky stiff, wherever on the net you may be
\documentclass[12pt]{article}
\usepackage{color}
\usepackage{hyperref}
\usepackage{mdwlist}
\usepackage{longtable}
\usepackage{ulem}
\usepackage{listings}
\usepackage{glossaries}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphics}

@@

\normalem
\hypersetup{colorlinks=true, linkcolor=blue, filecolor=blue, pagecolor=blue, urlcolor=blue}
\makeglossaries

@@

Creating an open implementation of a modern decompression algorithm could make diving much safer, and
provide divers with software that could be shown to be safe. This report outlines the design and implementation
of a proof of concept implementation of the VPM-B \newglossaryentry{VPM-B}{name={VPM-B}, description={Varying Permeability Model}}
algorithm, in the hopes that it can provide a base on 
which to build production quality software and hardware for an open dive computer.

@@

\subsubsection{VPM}
 The VPM algorithm starts by adopting a model of body tissues developed by Albert A. B\"uhlmann. This model
uses 16 tissue compartments to model the effect the dissolved gasses have on human tissues. During the
normal part of the dive, gas is added to the compartments\cite{Buhlmann84} (on-gassing\newglossaryentry{On-Gassing} {name={On-Gassing}, description={The process of adding inert gasses to body tissues}}
), and during the final ascent,
they're released from the tissue (off-gassing\newglossaryentry{Off-Gassing} {name={Off-Gassing}, description={The process of releasing inert gasses from body tissues}}
). Many different compartments are needed because different
tissues absorb and release gasses at different rates (as evidenced by the Helium\_Half\_Time and
Nitrogen\_Half\_Time arrays in the program).

@@

To reduce the time decompressing without increasing the danger, the Critical Volume Algorithm is applied
during the ascent. The CVA\newglossaryentry{CVA} {name={CVA}, description={Critical Volume Algorithm, the algorithm that allows the VPM algorithm to give decent decompression times.}}
can't reduce the time of a ``no bubble growth'' ascent, but it can limit
the growth of the bubbles to a safe amount. This is done by calculating a ``no bubble growth'' ascent,
and then applying the CVA iteratively until the ascent is judged safe.

@@

For this honours project, two open source implementations of the algorithm VPM-B algorithm were written. A version designed to be
usable in a desktop setting to serve as a reference implementation, and allow the user to generate decompression
tables, and a version able to work on an embedded computer, and do real time measurements of depth and time, to
build accurate models of the gasses in a divers system. Furthermore, these implementations are released under the
``Simplified BSD License'' \newglossaryentry{BSD License} {name={BSD License}, description={A family of software licenses that allow the use of the code in open source or proprietary software}}
, which allows them to be used in a proprietary or open source dive computer. While
this does allow a company to simply take the code, modify it, and use it in their product without
giving back the changes, it is an acceptable compromise to it not being used at all, because it is under
a more restrictive license.

@@

 Once the Python version was up and running, it was time to make many changes to the code to make it more
in line with the programming practices taught at Carleton University. The original program was written for FORTRAN 77,
and followed many of the practices common to FORTRAN\newglossaryentry{FORTRAN} {name={FORTRAN}, description={A programming language, popular with scientific researchers, but
not often taught to programmers these days}}
programs (GOTOs, global state used for all long term storage,
custom file formats for input data), which would be very hard for modern day programmers to follow. The program
was made object oriented\footnote{There are definite instances of the ``God Object'' anti-pattern, but at least the state is encapsulated, and in a position to be refactored}, by ensuring that all the global state was encapsulated in a DiveState object, and all functions that would modify this state belonged to the object.
The GOTOs were turned into normal loops with break statements, extremely long functions were broken into smaller
pieces, and repeating code was extracted to stand alone methods. Finally,  the input and output file formats were changed
to use JSON,\newglossaryentry{JSON} {name={JSON}, description={JavaScript Object Notation, a lightweight data representation language
designed to be easier to write than XML}}
 as it is a standard format like XML\newglossaryentry{XML} {name={XML}, description={Extensible Markup Language, a language used
to structure data for easy computer reading and writing}}
, that is supported by just about every programming language \footnote{Though not FORTRAN, so a translator was written to let the original program load the new input files}.

 Now that the Python program was finished, the port to C\newglossaryentry{C} {name={C}, description={A programming language, designed for small size, and fast execution, which is 
well suited for embedded systems}}
 was started. This was fairly straightforward, but it did involve
much more work to load JSON files than the Python version\footnote{The Python parser can turn it 
directly into hash-tables and arrays, while the C version
needs load values into the structures one at a time}. The port also showed a few bugs in the rarely run parts
of the Python program, that ended up being caught by the static type checks.
